The Privacy of Private Browsing

### 2013-12-15 finalproject/paper.tex
updated pdf

\hypersetup{
  colorlinks=false,
  pdfborder={0 0 0}
}
\setlength\parindent{0.5in}

@@

  \begin{abstract}
  Most modern web browsers have a ``private browsing'' mode that supposedly 
  allows a user to surf the internet without leaving any traces of his or her
  activity on his or her machine. However, the notion of ``private browsing'' 
  offers users a false sense of security, as browsing information is often left
  behind when a private browsing session terminates. Several researchers have 
  already demonstrated methods of detecting this information. These include the
  analysis of virtual and browser memory and the pagefile on Windows machines.
  The existence of private browsing artifacts on a user’s machine raises many 
  questions: how much privacy do web browser developers actually aim to provide
  with ``private browsing'' modes? Are these goals accurately conveyed to the 
  end user? And, if the user’s browsing session can be reconstructed from these
  artifacts, how might this be exploited by digital forensics professionals? 
  This paper seeks to answer these questions and to prove that such artifacts 
  can be used to determine, at least in part, what a user was doing during his 
  or her ``private browsing'' session---thereby rendering it not very private 
  at all.
  \end{abstract}

  \doublespacing
  \section{Introduction}
    \subsection{Browser Overview}
    Currently, the most popular browsers for the Windows family of operating 
    systems are Mozilla Firefox, Google Chrome, and Microsoft Internet 
    Explorer. All three of these browsers have built-in private browsing modes.
    The features and protections provided by each of these private browsing
    modes are summarized below.\cite{verdi13}\cite{google13}\cite{ie13}

    \singlespacing
    \begin{center}
      \begin{tabular}{|p{5cm}|c|c|c|}
        \hline
         & Firefox & Chrome & Internet Explorer \\
         & Private Browsing & Incognito Mode & InPrivate Browsng \\
        \hline
        Browser history & X & X & X \\
        \hline
        Form data & X &  & X \\
        \hline
        Search bar data & X &  & X \\
        \hline
        Passwords & X &  & X \\
        \hline
        Downloads & X & X &  \\
        \hline
        Cookies & X & X & X \\
        \hline
        Cached files & X &  & X \\
        \hline
        Simulataneous sessions with different privacy settings & X & X & X \\
        \hline
      \end{tabular}
    \end{center}

    \doublespacing
    \subsection{Definitions}
    Many shortcomings of private browsing modes relate to how a computer's 
    operating system manages memory and disk space. Those who are unfamiliar
    with these tasks may benefit from the definitions below.

    \begin{description}
      \item[Cluster] A group of disk sectors that the operating system
      treats as a unit; smallest unit of disk storage that the operating
      system can allocate
      \item[Digital certificate] Verifies that a particular website or target
      server is who it claims to be when connecting securely
      \item[Disk] The hard drive of a computer; where a user's files and
      directories (among other things) are stored
      \item[Free space] A cluster that is not allocated to any particular 
      file
      \item[Page] The smallest unit of memory that the operating system can
      allocate; can be temporarily stored on disk to allow the computer to run
      more programs at the same time
      \item[Slack space] The space between the end of a file and the end of 
      the cluster in which it is stored; important in computer forensics for
      finding fragments of old files\footnote{Hard drive clusters are of a 
      fixed size determined by the operating system. Most of the time, when a 
      user saves a file to his or her hard drive, the file is not exactly the 
      same size as the cluster in which the operating system puts it. 
      Furthermore, when a user ``deletes'' a file from his or her hard drive, 
      the file is not really deleted---it is simply marked as available to be
      overwritten if necessary. If a file is eventually stored on the hard 
      drive in a cluster that used to be home to a larger (now ``deleted'') 
      file, leftover data from the old file will still exist at the end of that
      cluster.}
    \end{description}

  \section{To the Community}
  There are many misconceptions about private browsing. Some believe that it
  prevents an internet service provider, network administrator, or attackers
  engaged in packet sniffing from linking a user's internet activities to his
  or her identity. It does not. Any and all network packets leaving the user's
  machine contain information such as the user's IP address, which can be used
  to determine his or her location and/or identity. Others think that private
  browsing sessions leave absolutely no trace of their activities on their
  local machines. As the next section of this paper will describe, this is also
  false. Still others think that private browsing will prevent the National
  Security Agency from tracking them on social media. It certainly does not.
  This paper will shed some light on what private browsing sessions can and
  cannot do, allowing internet users to think twice before exploring the web.

  \section{Applications: Forensics}
  One of the biggest problems with private browsing sessions is that artifacts
  remain on a machine after the user exits his or her internet browser. These
  artifacts can include logon information for websites, browsing history,
  and digital media such as images and videos---all items that you would expect
  a private browsing session to keep private.\cite{ohana13} Private browsing
  sessions can be exploited by browser extensions and remote attacks, but this 
  section will focus on artifacts left on a local machine by the browsers 
  themselves.

@@

>>>>

\hypersetup{
  colorlinks=false,
  pdfborder={0 0 0}
}
\setlength\parindent{0.5in}
\doublespacing

@@

  \begin{abstract}
  Most modern web browsers have a ``private browsing'' mode that supposedly 
  allows a user to surf the internet without leaving any traces of his or her
  activity on his or her machine. However, the notion of ``private browsing'' 
  offers users a false sense of security, as browsing information is often left
  behind when a private browsing session terminates. Several researchers have 
  already demonstrated methods of detecting this information. These include the
  analysis of the hard drive, memory, and the pagefile on Windows machines.
  The existence of private browsing artifacts on a user’s machine raises many 
  questions: how much privacy do web browser developers actually aim to provide
  with ``private browsing'' modes? Are these goals accurately conveyed to the 
  end user? And, if the user’s browsing session can be reconstructed from these 
  artifacts, how much information would digital forensics professionals be 
  able to gain? This paper seeks to answer these questions and to prove that 
  such artifactscan be used to determine, at least in part, what a user was 
  doing during his or her ``private browsing'' session---thereby rendering it 
  not very private at all.
  \end{abstract}

  \section{Introduction}
    \subsection{Browser Overview}
    Currently, the most popular browsers for the Windows family of operating 
    systems are Mozilla Firefox, Google Chrome, and Microsoft Internet 
    Explorer. All three of these browsers have built-in private browsing modes.
    While they all have different names---``Private Browsing'' in Firefox,
    ``Incognito Mode'' in Chrome, and ``InPrivate Browsing'' in Internet
    Explorer---they all claim to do essentially the same thing: allow a user to
    browse the web without traces of his or her browsing history, login
    information, form and search bar entries, or downloads remaining on his or
    her machine.\cite{verdi13}\cite{google13}\cite{ie13}

    \subsection{Definitions}
    Many shortcomings of private browsing modes relate to how a computer's 
    operating system manages memory and disk space. Those who are unfamiliar
    with these tasks may benefit from the definitions below.

    \singlespacing
    \begin{itemize}
      \item \emph{Cluster}: A group of disk sectors that the operating system
      treats as a unit; smallest unit of disk storage that the operating
      system can allocate
      \item \emph{Digital certificate}: Verifies that a particular website or 
      target server is who it claims to be when connecting securely
      \item \emph{Disk}: The hard drive of a computer; where a user's files and
      directories (among other things) are stored
      \item \emph{Free space}: A cluster that is not allocated to any 
      particular file
      \item \emph{Page}: The smallest unit of memory that the operating system 
      can allocate; can be temporarily stored on disk to allow the computer to 
      run more programs at the same time
      \item \emph{Slack space}: The space between the end of a file and the end
      of the cluster in which it is stored; important in computer forensics for
      finding fragments of old files\footnote{Hard drive clusters are of a 
      fixed size determined by the operating system. Most of the time, when a 
      user saves a file to his or her hard drive, the file is not exactly the 
      same size as the cluster in which the operating system puts it. 
      Furthermore, when a user ``deletes'' a file from his or her hard drive, 
      the file is not really deleted---it is simply marked as available to be
      overwritten if necessary. If a file is eventually stored on the hard 
      drive in a cluster that used to be home to a larger (now ``deleted'') 
      file, leftover data from the old file will still exist at the end of that
      cluster.}
    \end{itemize}

  \doublespacing
  \section{To the Community}
  There are many misconceptions about private browsing. Some believe that it
  prevents an internet service provider, network administrator, or attackers
  engaged in packet sniffing from linking a user's internet activities to his
  or her identity. It does not. Any and all network packets leaving the user's
  machine contain information such as the user's IP address, which can be used
  to determine his or her location and/or identity. Others think that private
  browsing sessions leave absolutely no trace of their activities on their
  local machines. As the next section of this paper will describe, this is also
  false. Still others think that private browsing will prevent the National
  Security Agency from tracking them on social media. It certainly does not.
  This paper will shed some light on what private browsing sessions can and
  cannot do, allowing internet users to think twice before exploring the web.

  \section{Applications: Forensics}
  One of the biggest problems with private browsing sessions is that artifacts
  remain on a machine after the user exits his or her internet browser. These
  artifacts can include logon information for websites, browsing history,
  and digital media such as images and videos---all items that you would expect
  a private browsing session to keep private. Additionally, browser extensions 
  can often record this kind of information, even when a user is browsing 
  privately.

@@

    %\subsection{Browser Extensions}
    % Write me!
